BVP Paper

### 2017-12-11 tex/bvp_paper.tex
updates paper and biblio a bit

Studies of convection in stratified systems which aim to model convection 
in astrophysical systems, such as the outer envelopes of low-to-moderate mass stars 
like the Sun, often employ a mixture of these
two types of boundary conditions \cite{hurlburt&all1984, cattaneo&all1991}.  
The flux at the lower boundary is fixed, modeling
the constant energy generation of the stellar core, 
while the outer boundary condition is held at a fixed temperature,
modeling the surface of a star which must output the energy generated internally.

@@

\subsection{Governing Equations, Nondimensionalization, and Domain setup}
In our first set of experiments, we adopt the Oberbeck-Boussinesq approximation.  Here, the
fluid has constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$).  We non-dimensionalize length by the layer height ($L_z$) ,
temperature by the (constant) initial temperature gradient across the layer ($\grad T_0$), and time
by the freefall timescale ($L_z / v_{\text{ff}}$, with $v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$, where $g$ is 
uniform gravitational acceleration in the $-\hat{z}$ direction). The dimensionless Boussinesq
equations governing the velocity $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$, temperature
$T = T_0 + T_1$, and reduced pressure $\varpi$ are (cite Rayleigh, chandrasekar)
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\grad\varpi + T_1\hat{z} + \frac{\text{Pr}}{\text{Ra}}\grad^2\bm{u}, 
	\label{eqn:bouss_momentum_eqn}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \frac{1}{\text{Pr}\,\text{Ra}}\grad^2 T_1,
	\label{eqn:bouss_energy_eqn}
\end{gather}
where the dimensionless control parameters are the Rayleigh and Prandtl numbers,
\begin{equation}
\text{Ra} = \frac{g \alpha L^4 \left(\frac{dT}{dz}\right)_0}{\nu\chi} = \frac{(L\,v_{\text{ff}})^2}{\nu\chi}, \qquad \text{Pr} = \frac{\nu}{\chi}.
\end{equation}
The dimensionless vertical extent of the domain is $z = [-1/2, 1/2]$, and at the boundaries
we impose no-slip, impenetrable boundary conditions such that $w = u = v = 0$ at $z = \pm 1/2$.
At the lower boundary, we employ a fixed flux condition such that $\partial T_1 / \partial z = 0$
at $z = -1/2$, and we impose a fixed temperature condition $T_1 = 0$ at $z = 1/2$. Both
horizontal directions are periodic and have equal aspect ratio, $\gamma = 4$, such that
the horizontal coordinates $x, y = [0, \Gamma]$.

@@

\subsection{Numerics}

@@

\section{Fully Compressible Convection}

\subsection{Governing Equations}
We study stratified convection in an ideal gas whose adiabatic index is $\gamma = 5/3$.
The initial atmospheric stratification is polytropic \cite{anders&brown2017}.
We assume a Newtonian radiative conduction term \cite{lecoanet&all2014}, and solve the
fully compressible Navier-Stokes equations of the form
\begin{gather}
\frac{\partial \ln \rho}{\partial t} + \bm{u}\cdot\grad\ln\rho + \DivU = 0
	\label{eqn:FC_continuity}
\\
\frac{D \bm{u}}{Dt} = -T\grad\ln\rho - \grad T + \bm{g} - \Div{\stressT}
	\label{eqn:FC_momentum}
\\
\frac{D T}{D t} + (\gamma-1)T\DivU = 
\frac{1}{\rho c_V}\left(\kappa\grad^2 T - [\stressT\cdot\grad]\cdot\bm{u}\right),
	\label{eqn:FC_energy}
\end{gather}
where $D/Dt \equiv \partial/\partial t + \bm{u}\cdot\grad$ and the viscous stress
tensor is defined as
\begin{equation}
\Pi_{ij} \equiv -\mu\left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij
}\DivU\right)
\label{eqn:stress_tensor}
\end{equation}
and $\delta_{ij}$ is the Kronecker delta function.

\subsection{The Boundary Value Equations}
In studies of fully compressible convection, the flux carried by the adiabatic temperature gradient
is not available for convection.  Thus, only the flux \emph{in excess} of the adiabat will drive
convection and be carried by convection.  As such, this is the only portion of the flux which
must be examined to determine if the solution is in a converged state.  In a system with
a fixed flux boundary condition, the available superadiabatic flux is
\begin{equation}
F_{\text{avail}} = -\kappa(\grad T_0 - \grad T_{\text{ad}}) = \kappa\frac{\epsilon}{c_P}\grad T_0,
\end{equation}
which is small when $\epsilon$ is small and only requires low Mach number convective flows
to carry it.  In a perfectly evolved solution, there will be thin boundary layers in which
conduction carries this flux in addition to the adiabatic flux, but in an efficient convective
interior, convective fluxes must carry this full amount.

The BVP equations are inspired by equations of stellar modeling  \cite{paxton&all2011} but adapted to these simulations
of fully compressible convection.  Here, rather than parameterizing convection, we can get the
convective fluxes directly from our simulation adn use them to solve for the appropriate structure
of the atmosphere.  The FC BVP equations are
\begin{gather}
\frac{d M_1}{dz} = \rho_1 \\
T_0 \grad \rho_1 + T_1 \grad \rho_0 + \rho_1 g = - T_0 \grad \rho_0 - T_1 \grad \rho_1 - \rho_0 g \\
\kappa \frac{d^2 T_1}{d z^2} = -\frac{d}{dz}F_{\text{conv, z}},
\end{gather}
which ensure mass conservation, thermal equilibrium, and that the atmosphere is, on average,
in hydrostatic equilibrium.  We couple these equations with four boundary conditions (mixed
flux / temperature boundary conditions, as well as setting $M_1 = 0$ at the top and bottom
of the atmosphere).

\subsection{Results}

>>>>

Studies of convection in stratified systems which aim to model convection 
in astrophysical systems, such as the outer envelopes of low-to-moderate mass stars 
like the Sun, often employ a mixture of these
two types of boundary conditions \cite{hurlburt&all1984, cattaneo&all1991}.  
The flux at the lower boundary is fixed, modeling
the constant energy generation of the stellar core, 
while the outer boundary condition is held at a fixed temperature,
modeling the surface of a star which must output the energy generated internally.
This set of boundary conditions is also being studied more recently in Boussinesq convection
\cite{korre&all2017}, and is displaying interesting characteristics such as internal layers which
are stable to convection.

@@

\subsection{Governing Equations, Nondimensionalization, and Domain setup}
In our first set of experiments, we adopt the Oberbeck-Boussinesq approximation.  Here, the
fluid has constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$).  We non-dimensionalize length by the layer height ($L_z$) ,
temperature by the (constant) initial temperature gradient across the layer ($\grad T_0$), and time
by the freefall timescale ($L_z / v_{\text{ff}}$, with $v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$, where $g$ is 
uniform gravitational acceleration in the $-\hat{z}$ direction). The dimensionless Boussinesq
equations governing the velocity $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$, temperature
$T = T_0 + T_1$, and reduced pressure $\varpi$ are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\grad\varpi + T_1\hat{z} + \frac{\text{Pr}}{\text{Ra}}\grad^2\bm{u}, 
	\label{eqn:bouss_momentum_eqn}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \frac{1}{\text{Pr}\,\text{Ra}}\grad^2 T_1,
	\label{eqn:bouss_energy_eqn}
\end{gather}
where the dimensionless control parameters are the Rayleigh and Prandtl numbers,
\begin{equation}
\text{Ra} = \frac{g \alpha L^4 \left(\frac{dT}{dz}\right)_0}{\nu\chi} = \frac{(L\,v_{\text{ff}})^2}{\nu\chi}, \qquad \text{Pr} = \frac{\nu}{\chi}.
\end{equation}
The dimensionless vertical extent of the domain is $z = [-1/2, 1/2]$, and at the boundaries
we impose no-slip, impenetrable boundary conditions such that $w = u = v = 0$ at $z = \pm 1/2$.
At the lower boundary, we employ a fixed flux condition such that $\partial T_1 / \partial z = 0$
at $z = -1/2$, and we impose a fixed temperature condition $T_1 = 0$ at $z = 1/2$. Both
horizontal directions are periodic and have equal aspect ratio, $\gamma = 4$, such that
the horizontal coordinates $x, y = [0, \Gamma]$.

@@

In essence, solving the BVP is equivalent to saying that even though the magnitude of flux through
the system is not initially correct, the system appropriately picks out the ratios
\begin{equation}
f_{\text{conv}} = \frac{F_{\text{conv}}}{F_{\text{tot}}}\qquad
f_{\text{cond}} = \frac{F_{\text{cond}}}{F_{\text{tot}}}.
\end{equation}
in the convective, transient state, but the total amount of flux is too large until the system
reaches the proper isotherm.

\subsection{Numerics}
Copy+paste from last paper:

We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Thermodynamic variables are decomposed such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, 
and the velocity is $\bm{u} = w\bm{\hat{z}} + u\bm{\hat{x}} + v\bm{\hat{y}}$.
In our 2D runs, $v = 0$.
Subscript 0 variables, set by (\ref{eqn:polytrope}), 
have no time derivative and vary only in $z$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Domain sizes range from
64x256 coefficients at the lowest values of 
Ra to 1024x4096 coefficients at Ra $> 10^{7}$ in 2D,
and from 64x128$^2$ to  256x512$^2$ in 3D. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at high ($\sim 1$) 
and low ($\sim 10^{-4}$) Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

We impose impenetrable, stress free, fixed temperature boundary conditions at
the top and bottom of the domain, with 
$w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. 
$T_1$ is initially filled with
random white noise whose magnitude is infinitesimal
compared to $\epsilon T_0$.
We filter this noise spectrum in coefficient space, 
such that only the lower 25\% of the coefficients
have power. All reported results are taken from time averages
over many $t_b$ beginning \{100, 40\}$t_b$
after the start of our \{2D, 3D\} simulations to
ensure our results are not biased by the convective transient.

@@
